---
title: "Analysis of DataCo Supply chain"
author: "Ibrahim"
date: "2023-04-11"
output:
  word_document: default
  html_document: default
---


#Load the required libraries
```{r}
library(ggplot2)
library(dplyr)
library(tidyr)
library(RColorBrewer)
library(plotrix)
library(lattice)
```

#Read CSV file and load it in a data frame
We can see top elements of the data
```{r}
#Reading the dataset
supply_chain = read.csv('C:/Users/Lenovo/OneDrive/Documents/directory files/DataCoSupplyChainDataset.csv')

head(supply_chain)
#View(supply_chain)
```

### Check the structure of the data frame
We can see that the data is mix of Int and Chars.
```{r}
str(supply_chain)
colnames(supply_chain)  # column names
```

### checking NA values
There are 336201 NA values in Product Description and 
Order Zipcode columns.
```{r}
colSums(is.na(supply_chain))
sum(is.na(supply_chain))

```
#Deleting null values
```{r}
supply_chain <- supply_chain[, -which(names(supply_chain) %in% c("Order.Zipcode", "Product.Description"))]
#supply_chain
```

#Checking null values
```{r}
#colSums(is.na(supply_chain))
sum(is.na(supply_chain))
```

#Order Item Discount Distribution
```{r}
ggplot(supply_chain, aes(x = Order.Item.Discount)) +
  geom_histogram(binwidth = 1, fill = "blue", color = "black") +
  labs(title = "Order Item Discount Distribution",
       x = "Order Item Discount",
       y = "Count")
```

#Count of mode of transaction
```{r}
type_count <- table(supply_chain$Type)
type_count
```

#Pie chart on Count of mode of transaction
```{r}
pie(type_count)
title(main = "Count of Mode of Transaction" )
```

#Taking only 10000 values to generate graphs quickly
```{r}
supply_chain_sample <- sample_n(supply_chain, 1000)
#supply_chain_sample
```

#Calculating Avg. number of days to deliver
```{r}
mean(supply_chain$Days.for.shipping..real.)  #average time to delivery is 3-4 days
```

#Showing average time delivery in a plot
```{r}
bwplot(~Days.for.shipping..real. , data=supply_chain)
#title(main = "Avg. number of days to deliver")
```

#Bar graph on Order Status
```{r}
ggplot(data = supply_chain, aes(x = Order.Status)) +
  geom_histogram(stat = "count", fill = "steelblue", color = "black") +
  ggtitle("Order Status") +
  xlab("Order Status") +
  ylab("Frequency")
```
#Number of Orders from different product categories between shipping modes
```{r}
# Subset data for each shipping mode
first_class_department <- subset(supply_chain, Shipping.Mode == "First Class")
second_class_department <- subset(supply_chain, Shipping.Mode == "Second Class")
standard_class_department <- subset(supply_chain, Shipping.Mode == "Standard Class")
same_day_department <- subset(supply_chain, Shipping.Mode == "Same Day")

# Aggregate data for each department
department <- aggregate(cbind(number = Delivery.Status) ~ Department.Name, data = supply_chain, FUN = function(x) length(x))
first_class_department <- aggregate(cbind(number = Delivery.Status) ~ Department.Name, data = first_class_department, FUN = function(x) length(x))
second_class_department <- aggregate(cbind(number = Delivery.Status) ~ Department.Name, data = second_class_department, FUN = function(x) length(x))
standard_class_department <- aggregate(cbind(number = Delivery.Status) ~ Department.Name, data = standard_class_department, FUN = function(x) length(x))
same_day_department <- aggregate(cbind(number = Delivery.Status) ~ Department.Name, data = same_day_department, FUN = function(x) length(x))

# Create plot
fig <- ggplot() +
  geom_line(data = department, aes(x = Department.Name, y = number, group = 1), size = 1) +
  labs(title = "Total number of orders from different product categories",
       x = "Department Name",
       y = "Number of Orders") +
  theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
  facet_wrap(~ "", nrow = 1)

# Example of using group aesthetic in geom_line()
fig2 <- ggplot() +
  geom_line(data = first_class_department, aes(x = Department.Name, y = number, color = "First Class", group = 1), size = 1) +
  geom_line(data = second_class_department, aes(x = Department.Name, y = number, color = "Second Class", group = 1), size = 1) +
  geom_line(data = standard_class_department, aes(x = Department.Name, y = number, color = "Standard Class", group = 1), size = 1) +
  geom_line(data = same_day_department, aes(x = Department.Name, y = number, color = "Same Day", group = 1), size = 1) +
  labs(title = "Number of Orders from different product categories between shipping modes",
       x = "Department Name",
       y = "Number of Orders",
       color = "Shipping Mode") +
  theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
  facet_wrap(~ "", nrow = 1) +
  scale_color_manual(values = c("First Class" = "blue", "Second Class" = "green", "Standard Class" = "red", "Same Day" = "purple"))

print(fig2)

# Display plots
print(fig)
#print(fig2)
```

```{r}
summary(supply_chain)
```


